---
title: "Review"
format:
  revealjs:
    slide-number: true
    preview-links: true
    scrollable: true
---

## Exercise: Data Frame Normalization Using sweep

Normalize a data frame df (with columns `X1`, `X2`, `X3` each containing $10$ random integers between $1$ and $100$) by subtracting the median and dividing by the interquartile range of each column.

::: {.fragment}

### Creating a data frame

```{r echo=TRUE,eval=TRUE}
set.seed(123)
df = data.frame(X1 = sample(1:100, 10), 
                X2 = sample(1:100, 10), 
                X3 = sample(1:100, 10))

df
```


### Calculating medians and interquartile ranges
```{r echo=TRUE,eval=TRUE}
medians = sapply(df, median)
medians
iqrs = sapply(df, IQR)
iqrs
```

### Normalizing the data frame
```{r echo=TRUE,eval=TRUE}
normalized_df = sweep(sweep(df, 2, medians, "-"), 2, iqrs, "/")
normalized_df
```

:::


## Exercise: Generating and Analyzing Height Data

Generate a dataset that simulates the heights (in centimeters) of 1000 individuals. Assume an average height of 170 cm and a standard deviation of 10 cm. Follow the following steps:

- Create the dataset using a normal distribution with the given mean and standard deviation using the `rnorm()` function.
- Calculate the $Z$-scores for the entire dataset using the `sweep()` function.
- Determine how many standard deviations away from the mean is a height of 185 cm.
